Elaboration with First-Class Implicit Function Types

### 2020-05-10 paper.tex
revs

Secondly, FCIF supports global inference. For an example for global inference,
$\slet\,x : \_ = \mi{single}\,\mi{id}\,\sin\,$
$\mi{cons}\,\{\mi{IdTy}\}\,x\,\mi{nil}$ works in FCIF, MLF and HML but does not
work in GI, QL and HMF. MLF and HML work here by immediately giving a principal
type to $single\,id$ which involves generalization with a subtyping bound. In
contrast, FCIF makes no promises about principal typing --- which is not
feasible with dependent types --- and does no generalization, but it can still
infer a type for $id$ which can be later constrained to $IdTy$.

\subsection{Inferring Polymorphic Types for Function Arguments}
\label{sec:appinsert}

>>>>

Secondly, FCIF supports global inference. For an example for global inference,
$\slet\,x : \_ = \mi{single}\,\mi{id}\,\sin\,$
$\mi{cons}\,\{\mi{IdTy}\}\,x\,\mi{nil}$ works in FCIF, MLF and HML but does not
work in GI, QL and HMF. MLF and HML work here by immediately giving a principal
type to $\mi{single}\,\mi{id}$ which involves generalization with a subtyping bound. In
contrast, FCIF makes no promises about principal typing --- which is not
feasible with dependent types --- and does no generalization, but it can still
infer a type for $\mi{id}$ which can be later constrained to $\mi{IdTy}$.

\subsection{Polymorphic Argument Inference and Curried Application Insertion}
\label{sec:appinsert}

FCIF cannot infer the $B1$ and $B2$ cases on Figure \ref{fig:benchmark}, which
would require inferring polymorphic types for function arguments. In very simple
cases, FCIF is able to infer (ambiguous) polymorphic argument types. For
example, it can infer a type for $(\lambda\,f.\,f\{\mi{zero}\})$. Assuming
$\mi{Pair} : \U \to \U \to \U$ and $\mi{pair}: \{A\,B : \U\}\to A \to B \to
\mi{Pair}\,A\,B$, it can also infer type for the following:
\[
    \lambda\,f\,x.\,\mi{pair}\,(f\,\{x\}\,x)\,(f\,x)
\]
Here, FCIF can proceed because it first encounters $(f\,\{x\}\,x)$, where it
learns that $f$ should have two arguments, where the first one is implicit.
Armed with this knowledge, it inserts a fresh meta in $(f\,x)$. The following
fails though:
\[
    \lambda\,f\,x.\,\mi{pair}\,(f\,x)\,(f\,\{x\}\,x)
\]
Now, FCIF decides upon elaborating $(f\,x)$ that $f$ is a function whose first
argument is explicit, and fails at $f\,\{x\}$. This is similar to the implicit
$\lambda$ insertion issue, but here we would have to refine implicit
applications. We could try inserting an unknown curried application after the
first $f$, which could be later refined. We sketch an extension of elaboration
which could possibly handle this example. We extend insertion with
the case which inserts a new curried application whenever we have a meta-headed
type. We omit metacontexts and metasubstitutions for brevity.
\begin{alignat*}{3}
& \einsert\,(t,\,\alpha\,\ol{u}) :\equiv \edo \\
& \quad A \leftarrow \mathsf{freshMeta}_{\Gamma}\,\Tel\\
& \quad B \leftarrow \mathsf{freshMeta}_{\Gamma,\,x: \Rec\,A}\,\U\\
& \quad \unify\,(\alpha\,\ol{u})\,(\{x : \ol{A}\} \to B)\\
& \quad \mathsf{newConstancy}_{\Gamma, x : \Rec\,A}\,B\\
& \quad u \leftarrow \mathsf{freshMeta}_{\Gamma}\,(\Rec\,A)\\
& \quad \ereturn\,(t\,\{u : \ol{A}\},\,B[x \mapsto u])
\end{alignat*}

%% let e2 = λ f (x : Int). pair (f {id {Int} x} x) (f x) in
